The Next Generation of Canadian Salmon

### 2024-04-19 output/paper.qmd
add remaining citations

The declining trend in Canadian salmon runs, as observed by many avid anglers and supported by empirical data (INSERT CITATION), serves as the focal point of this paper. Once thriving waterways have experienced a significant reduction in salmon populations, posing challenges not only for recreational anglers but also for the fishing industry, a vital component of Canadian exports and numerous livelihoods, particularly in coastal regions. Due to this significance to many Canadians, it is important to investigate the decline in salmon runs, highlighting the importance of understanding and addressing these challenges that are faced in the effort to safeguard future salmon generations.

@@

For information regarding the populations of Canadian Salmon in the Pacific region, the *New Salmon Escapement Database System (NuSEDS)* was used to retrieve the quality and health of the Salmon population. This data is not only used to measure the populations but other health and success measures that certain area salmon have. These observations are made during the spawning season for salmon (from September - November), recording the amount of salmon that have come to spawn in a certain waterbody. As it is not possible for those wildlife officials to individually count each salmon that enters a waterway at any time during the day, an estimated count was made based on the abundance of salmon currently present during the observation.

@@

For information regarding the salmon caught in commercial fisheries, the *Pacific Region Commercial Salmon Fishery In-season Catch Estimates* was used. This data comes from the Fishery Operations System (FOS) which serves as the primary database for Fisheries and Oceans Canada's (DFO) Pacific Region Commercial Salmon Logbook program, established in 1998 and mandated fleet-wide participation by 2001. This program compels commercial salmon fishers to log daily catch and effort data in harvest logbooks, subsequently reporting it to DFO via service providers within specified deadlines. However, certain personal information provided by fishers is protected and inaccessible. Fishery Managers utilize this reported data, alongside additional sources like overflights, to compute in-season catch estimates. Its important to note that this dataset omits data from test fishing, recreational fishing, or First Nations harvest.

>>>>

The declining trend in Canadian salmon runs, as observed by many avid anglers and supported by empirical data [@citedecline], serves as the focal point of this paper. Once thriving waterways have experienced a significant reduction in salmon populations, posing challenges not only for recreational anglers but also for the fishing industry, a vital component of Canadian exports and numerous livelihoods, particularly in coastal regions. Due to this significance to many Canadians, it is important to investigate the decline in salmon runs, highlighting the importance of understanding and addressing these challenges that are faced in the effort to safeguard future salmon generations.

@@

For information regarding the populations of Canadian Salmon in the Pacific region, the *New Salmon Escapement Database System (NuSEDS)* [@citenuseds] was used to retrieve the quality and health of the Salmon population. This data is not only used to measure the populations but other health and success measures that certain area salmon have. These observations are made during the spawning season for salmon (from September - November), recording the amount of salmon that have come to spawn in a certain waterbody. As it is not possible for those wildlife officials to individually count each salmon that enters a waterway at any time during the day, an estimated count was made based on the abundance of salmon currently present during the observation.

@@

For information regarding the salmon caught in commercial fisheries, the *Pacific Region Commercial Salmon Fishery In-season Catch Estimates* [@citecatch] was used. This data comes from the Fishery Operations System (FOS) which serves as the primary database for Fisheries and Oceans Canada's (DFO) Pacific Region Commercial Salmon Logbook program, established in 1998 and mandated fleet-wide participation by 2001. This program compels commercial salmon fishers to log daily catch and effort data in harvest logbooks, subsequently reporting it to DFO via service providers within specified deadlines. However, certain personal information provided by fishers is protected and inaccessible. Fishery Managers utilize this reported data, alongside additional sources like overflights, to compute in-season catch estimates. Its important to note that this dataset omits data from test fishing, recreational fishing, or First Nations harvest.
